---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(magrittr)
library(stringr)
library(targets)
```

# Overview

A data science project built with [quarto](https://quarto.org/) and [targets](https://books.ropensci.org/targets/).

# Setup

-   Create an R script called `_targets_config.R` in the project root directory with the following objects and populate:

```{r results='asis', eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
result <- readLines(tar_read(TARGETS_CONFIG)) %>%
  # subset for comments, object assignments and empty lines
  subset(.,
         str_detect(.,
                    "^#+ |<-|^$")) %>%

  # remove assignments, leaving only object names
  str_replace("<-.*$",
              "<-") %>%

  # print
  paste(sep = "",
        collapse = "\n")

cat(paste0("```",
           "\n",
           result,
           "\n",
           "```"))
```

-   Create a `.Renviron` file in the project root directory, specifying the paths to your local copy of `all_lkps_maps.db` and your targets config file (if using a different file) e.g. 

```
ALL_LKPS_MAPS_DB=~/Documents/Data/all_lkps_maps_sct.db
TARGETS_CONFIG=_targets_config_alternative.R
```

-   Install required R packages, ideally using [renv](https://rstudio.github.io/renv/) with `renv::init()`, or `renv::restore()` if a `renv.lock` file is already present.
-   Run [targets pipeline](https://books.ropensci.org/targets/) with `tar_make()` in the R console.
